---
title: "Generate Top Lists"
output: html_notebook
---


# Disclaimer
This document describes how results of predicted breeding values are read from `.csv` files for the two breeds `BV` and `OB`. The results read are reformatted into a series of tibbles which eventually are converted into an `xlsx` workbook. 



# Reading The Data
The data is the result of the breeding value prediction process. An example dataset is available in this project under the directory `inst/extdata`. Let us have a look at the available input files first.

```{r}
s_data_dir <- file.path(here::here(), "inst/extdata")
(vec_data_files <- list.files(s_data_dir, full.names = TRUE))
```

We start by reading the first data file into a tibble

```{r}
tbl_tlbv <- readr::read_csv2(file = vec_data_files[1])
tbl_tlbv
```



# Extracting the Top-List for One Trait
Columns of `tbl_tlbv` contain the following values

```{r}
(vec_all_cols <- colnames(tbl_tlbv))
```

We define the set of traits to be all of the column names except `Tiername`, `TVD-NR` and `Anbieter`. 

```{r}
vec_non_trait <- c("Tiername", "TVD-Nr", "Anbieter")
(vec_traits <- setdiff(vec_all_cols, vec_non_trait))
```

For each trait, we extract the top $N$ entries and list a pre-defined set of columns.

```{r}
(vec_result_cols <- setdiff(vec_non_trait, "Anbieter"))
```

The top animals are listed according to the values given in the trait column

```{r}
n_nr_top <- 12
tbl_tlbv[order(tbl_tlbv$GZW, decreasing = TRUE), c(vec_result_cols, "GZW")][1:n_nr_top,]
```



# Generalize Over All Traits
To be able to generalize the above approach, we have to use the trait name as a variable. 

```{r}
s_trait_name <- "GZW"
```

Then the extraction has to be changed slightly. In addition to the result extraction, ranks are added as the left most column using `bind_cols()`

```{r}
dplyr::bind_cols(tibble::tibble(Rang = 1:n_nr_top), tbl_tlbv[order(tbl_tlbv[[s_trait_name]], decreasing = TRUE), c(vec_result_cols, s_trait_name)][1:n_nr_top,])
```

The same extraction is done for every trait using `lapply()`.

```{r}
l_result_tl <- lapply(vec_traits, function(x) dplyr::bind_cols(tibble::tibble(Rang = 1:n_nr_top), 
                                                               tbl_tlbv[order(tbl_tlbv[[x]], decreasing = TRUE), c(vec_result_cols, x)][1:n_nr_top,]))
names(l_result_tl) <- vec_traits
l_result_tl
```



# Combine Everything Into a Single Function
We create a function that takes as input a list of breeds and associated input files. Furthermore, we have to specify the vector of traits and the vector of columns to appear in the result table. With that all information is extracted from the input files into a single list.

```{r}
read_top_list_info <- function(pl_breed_input, pvec_trait, pvec_resultcols){
  ### # output some parameter info
  cat("[INFO -- read_top_list_info] breeds:\n")
  print(pl_breed_input$breeds)
  cat("[INFO -- read_top_list_info] input files:\n")
  print(pl_breed_input$inputfiles)
  cat("[INFO -- read_top_list_info] number of top candidates:\n")
  print(pl_breed_input$numbertop)
  ### # initialize final result
  l_final_result <- NULL
  ### # loop over breeds and do the extraction
  for (nbidx in seq_along(pl_breed_input$breeds)){
    ### # read current input file
    tbl_cur_tl <- readr::read_csv2(file = pl_breed_input$inputfiles[nbidx])
    ### # extract the current result
    l_cur_result <- lapply(pvec_trait, function(x) {
                                       if (x %in% names(tbl_cur_tl)){
                                         return(dplyr::bind_cols(tibble::tibble(Rang = 1:pl_breed_input$numbertop[nbidx]), 
                                             tbl_cur_tl[order(tbl_cur_tl[[x]], decreasing = TRUE), c(pvec_resultcols, x)][1:pl_breed_input$numbertop[nbidx],]))
                                       } else {
                                         return(NULL)
                                       }})
    names(l_cur_result) <- pvec_trait
    ### # add current result to final result
    l_final_result <- c(l_final_result, list(l_cur_result))
  }
  ### # use the breeds as names for the final result list
  names(l_final_result) <- pl_breed_input$breeds
  ### # return list of final results
  return(l_final_result)
}
```

The above created function is tested with the following call. 

```{r}
vec_breed <- c("BV", "OB")
l_breed_input <- list(breeds = vec_breed,
                      inputfiles = vec_data_files,
                      numbertop = c(12, 5))
l_final_tl <- read_top_list_info(pl_breed_input = l_breed_input, pvec_trait = vec_traits, pvec_resultcols = vec_result_cols)
```


# Refinement of Trait Selection
The above lists which shows both breeds is a step in the right direction. We still have a problem that the selection of traits is only based on one breed. Hence we have to refine the seleciton of traits such that traits that are unique to one of the two breeds are also shown in the final result list.

The function `get_vec_traits()` extracts the trait names from the input files. 

```{r}
get_vec_traits <- function(pvec_input, pvec_non_trait){
  ### # initialize vector of trait names
  vec_trait_result <- NULL
  ### # read the first line of each input file
  for (ifidx in seq_along(pvec_input)){
    s_first_line <- readLines(pvec_input[ifidx], n = 1L)
    vec_cur_header <- unlist(strsplit(s_first_line, split = ";", fixed = TRUE))
    vec_trait_result <- union(vec_trait_result, setdiff(vec_cur_header, pvec_non_trait))
  }
  ### # return the final vector of traits
  return(vec_trait_result)
}
```


This leads to the following selection of traits. 

```{r}
(vec_det_traits <- get_vec_traits(pvec_input = l_breed_input$inputfiles, pvec_non_trait = vec_non_trait))
```

The toplists for both breeds can be re-generated as follows

```{r}
l_final_tl <- read_top_list_info(pl_breed_input = l_breed_input, pvec_trait = vec_det_traits, pvec_resultcols = vec_result_cols)
l_final_tl
```


# Output into xlsx
The top lists for both breeds is written to an excel workbook using `openxlsx`. 


```{r}
### # check availability of package and load it
if (!require(openxlsx)) install.packages('openxlsx')
library(openxlsx)
### # create a new workbook
wb <- createWorkbook()
### # add a table for the first trait
for (ctridx in seq_along(vec_det_traits)){
  s_cur_trait <- vec_det_traits[ctridx]
  cat(" * Current trait: ", s_cur_trait, "\n")
  addWorksheet(wb, s_cur_trait)
  ### # writing the data, start with trait name
  n_cur_start_row <- 2
  writeData(wb = wb, sheet = s_cur_trait, s_cur_trait, startCol = 1, startRow = n_cur_start_row)
  ### # increment start row
  n_cur_start_row <- n_cur_start_row + 1
  ### # write first block of data for first breed
  for (cbidx in seq_along(vec_breed)){
    s_cur_breed <- vec_breed[cbidx]
    cat(" ** Current breed: ", s_cur_breed, "\n")
    tbl_cur_tab <- l_final_tl[[s_cur_breed]][[s_cur_trait]]
    if (!is.null(tbl_cur_tab)){
      cat("    ==> tibble found\n")
      ### # write data frame for current breed
      writeData(wb = wb, sheet = s_cur_trait, tbl_cur_tab, startCol = 1, startRow = n_cur_start_row)
      ### # if we are not writing the last breed and the next breed also has toplist info, increment start row for second breed
      if (cbidx < length(vec_breed) && !is.null(l_final_tl[[vec_breed[cbidx+1]]][[s_cur_trait]])){
        n_cur_start_row <- n_cur_start_row + nrow(tbl_cur_tab) + 3
        writeData(wb = wb, sheet = s_cur_trait, vec_breed[cbidx + 1], startCol = 1, startRow = n_cur_start_row)
        n_cur_start_row <- n_cur_start_row + 2
      }
    }
  }
}

## Save workbook to working directory
saveWorkbook(wb, file = "Toplisten_Stiere_CHbv_1904.xlsx", overwrite = TRUE)
```


# Feature Request With Extended Column Names
Column names of the single tables should not be just the shown abbreviations, but they should be the real names. Hence we use a list that maps the abbreviations to the real names


```{r}
vec_det_traits
```

```{r}
l_trait_name <- lapply(vec_det_traits, function(x) return(x))
names(l_trait_name) <- vec_det_traits
```

Required adaptations are:

```{r}
l_trait_name$`ZW Milch` <- "Milch kg"
l_trait_name$EwP <- "Eiweiss %"
l_trait_name$ND <- "Nutzungsdauer"
l_trait_name$Per <- "Persistenz"
l_trait_name$FBK <- "Fruchtbarkeit"
l_trait_name$LBE <- "Gesamtnote"
l_trait_name$EU  <- "Euter"
l_trait_name$ZZ  <- "Zellzahl"
l_trait_name$RA  <- "Rahmen"
l_trait_name$BE  <- "Becken"
l_trait_name$FU  <- "Fundament"
```


These explicit trait names are used as additional arguments in the function that creates the lists with the information about the top-ranking animals. For reasons of simplicity, this function is re-defined here.

```{r}
read_top_list_info2 <- function(pl_breed_input, pl_trait, pvec_resultcols){
  ### # output some parameter info
  cat("[INFO -- read_top_list_info2] breeds:\n")
  print(pl_breed_input$breeds)
  cat("[INFO -- read_top_list_info2] input files:\n")
  print(pl_breed_input$inputfiles)
  cat("[INFO -- read_top_list_info2] number of top candidates:\n")
  print(pl_breed_input$numbertop)
  cat("[INFO -- read_top_list_info2] tratis: \n")
  vec_trait <- names(pl_trait)
  print(vec_trait)
  ### # initialize final result
  l_final_result <- NULL
  ### # loop over breeds and do the extraction
  for (nbidx in seq_along(pl_breed_input$breeds)){
    ### # read current input file
    tbl_cur_tl <- readr::read_csv2(file = pl_breed_input$inputfiles[nbidx])
    ### # extract the current result
    l_cur_result <- lapply(vec_trait, function(x) {
                                       if (x %in% names(tbl_cur_tl)){
                                         tbl_cur_trait <- dplyr::bind_cols(tibble::tibble(Rang = 1:pl_breed_input$numbertop[nbidx]), 
                                             tbl_cur_tl[order(tbl_cur_tl[[x]], decreasing = TRUE), c(pvec_resultcols, x)][1:pl_breed_input$numbertop[nbidx],])
                                         names(tbl_cur_trait)[ncol(tbl_cur_trait)] <- pl_trait[[x]]
                                         return(tbl_cur_trait)
                                       } else {
                                         return(NULL)
                                       }})
    names(l_cur_result) <- vec_trait
    ### # add current result to final result
    l_final_result <- c(l_final_result, list(l_cur_result))
  }
  ### # use the breeds as names for the final result list
  names(l_final_result) <- pl_breed_input$breeds
  ### # return list of final results
  return(l_final_result)
}
```

Testing the modified extraction function leads to 

```{r}
l_final_tl2 <- read_top_list_info2(pl_breed_input = l_breed_input, pl_trait = l_trait_name, pvec_resultcols = vec_result_cols)
l_final_tl2
```

## Third version of `read_top_list_info()`
Version 3 of the function reading the top-list information uses a `tibble` to map trait abbreviations to trait names. This makes it simpler to read this mapping from a file.

```{r}
read_top_list_info3 <- function(pl_breed_input, ptbl_trait, pvec_resultcols){
  ### # output some parameter info
  cat("[INFO -- read_top_list_info2] breeds:\n")
  print(pl_breed_input$breeds)
  cat("[INFO -- read_top_list_info2] input files:\n")
  print(pl_breed_input$inputfiles)
  cat("[INFO -- read_top_list_info2] number of top candidates:\n")
  print(pl_breed_input$numbertop)
  cat("[INFO -- read_top_list_info2] tratis: \n")
  vec_trait <- ptbl_trait$Abk
  print(vec_trait)
  ### # initialize final result
  l_final_result <- NULL
  ### # loop over breeds and do the extraction
  for (nbidx in seq_along(pl_breed_input$breeds)){
    ### # read current input file
    tbl_cur_tl <- readr::read_csv2(file = pl_breed_input$inputfiles[nbidx])
    ### # extract the current result
    l_cur_result <- lapply(vec_trait, function(x) {
                                       if (x %in% names(tbl_cur_tl)){
                                         tbl_cur_trait <- dplyr::bind_cols(tibble::tibble(Rang = 1:pl_breed_input$numbertop[nbidx]), 
                                             tbl_cur_tl[order(tbl_cur_tl[[x]], decreasing = TRUE), c(pvec_resultcols, x)][1:pl_breed_input$numbertop[nbidx],])
                                         names(tbl_cur_trait)[ncol(tbl_cur_trait)] <- ptbl_trait[ptbl_trait$Abk == x,]$Name
                                         return(tbl_cur_trait)
                                       } else {
                                         return(NULL)
                                       }})
    names(l_cur_result) <- vec_trait
    ### # add current result to final result
    l_final_result <- c(l_final_result, list(l_cur_result))
  }
  ### # use the breeds as names for the final result list
  names(l_final_result) <- pl_breed_input$breeds
  ### # return list of final results
  return(l_final_result)
}
```


Instead of a list, we can also use a tibble which can also be read from an input file

```{r}
tbl_trait_name <- tibble::tibble(Abk = vec_det_traits,
                                 Name = vec_det_traits)
```

Adaptations are done with the following statements

```{r}
tbl_trait_name[tbl_trait_name$Abk == "ZW Milch","Name"] <- "Milch kg"
tbl_trait_name[tbl_trait_name$Abk == "EwP", "Name"] <- "Eiweiss %"
tbl_trait_name[tbl_trait_name$Abk == "ND", "Name"]  <- "Nutzungsdauer"
tbl_trait_name[tbl_trait_name$Abk == "Per", "Name"] <- "Persistenz"
tbl_trait_name[tbl_trait_name$Abk == "FBK", "Name"] <-"Fruchtbarkeit"
tbl_trait_name[tbl_trait_name$Abk == "LBE", "Name"] <- "Gesamtnote"
tbl_trait_name[tbl_trait_name$Abk == "EU", "Name"]  <-  "Euter"
tbl_trait_name[tbl_trait_name$Abk == "ZZ", "Name"]  <-  "Zellzahl"
tbl_trait_name[tbl_trait_name$Abk == "RA", "Name"]  <-  "Rahmen"
tbl_trait_name[tbl_trait_name$Abk == "BE", "Name"]  <-  "Becken"
tbl_trait_name[tbl_trait_name$Abk == "FU", "Name"]  <-  "Fundament"  
tbl_trait_name
```

For future use, we write that `tibble` to a file

```{r}
s_trait_name_file <- "tl_trait_names.csv"
if (!file.exists(s_trait_name_file))
  readr::write_csv(tbl_trait_name, path = s_trait_name_file)
```

The trait names are now stored in `tbl_trait_name`. The access of the trait name via its abbreviation is different now. 

```{r}
tbl_trait_name[tbl_trait_name$Abk == "LBE", ][["Name"]]
tbl_trait_name[tbl_trait_name$Abk == "LBE", ]$Name
```

The vector with all trait abbreviations is obtained via

```{r}
tbl_trait_name$Abk
```

The toplist information is obtained via 

```{r}
l_final_tl3 <- read_top_list_info3(pl_breed_input = l_breed_input, ptbl_trait = tbl_trait_name, pvec_resultcols = vec_result_cols)
l_final_tl3
```


## Output Into xlsx Using A Function
Given the toplist information consisting of a list of lists of tibbles, we create a function that writes everything into an xlsx file. 

```{r}
write_tl_xlsx <- function(pl_toplist, ptbl_trait, ps_xlsx_file = "example.xlsx"){
  ### # extract trait abbreviations
  vec_det_traits <- ptbl_trait$Abk
  ### # vector of the breeds is taken from names of pl_toplist
  vec_breed <- names(pl_toplist)
  ### # create a new workbook
  wb <- openxlsx::createWorkbook()
  ### # add a table for the first trait
  for (ctridx in seq_along(vec_det_traits)){
    s_cur_trait <- vec_det_traits[ctridx]
    cat(" * Current trait: ", s_cur_trait, "\n")
    addWorksheet(wb, s_cur_trait)
    ### # writing the data, start with trait name
    n_cur_start_row <- 2
    writeData(wb = wb, sheet = s_cur_trait, s_cur_trait, startCol = 1, startRow = n_cur_start_row)
    ### # increment start row
    n_cur_start_row <- n_cur_start_row + 1
    ### # write first block of data for first breed
    for (cbidx in seq_along(vec_breed)){
      s_cur_breed <- vec_breed[cbidx]
      cat(" ** Current breed: ", s_cur_breed, "\n")
      tbl_cur_tab <- pl_toplist[[s_cur_breed]][[s_cur_trait]]
      if (!is.null(tbl_cur_tab)){
        cat("    ==> tibble found\n")
        ### # write data frame for current breed
        writeData(wb = wb, sheet = s_cur_trait, tbl_cur_tab, startCol = 1, startRow = n_cur_start_row)
        ### # if we are not writing the last breed and the next breed also has toplist info, increment start row for second breed
        if (cbidx < length(vec_breed) && !is.null(pl_toplist[[vec_breed[cbidx+1]]][[s_cur_trait]])){
          n_cur_start_row <- n_cur_start_row + nrow(tbl_cur_tab) + 3
          writeData(wb = wb, sheet = s_cur_trait, vec_breed[cbidx + 1], startCol = 1, startRow = n_cur_start_row)
          n_cur_start_row <- n_cur_start_row + 2
        }
      }
    }
  }
  
  ## Save workbook to working directory
  saveWorkbook(wb, file = ps_xlsx_file, overwrite = TRUE)
  ### # return nothing
  return(invisible(TRUE))
}
```

How can we determine the vector of the breeds

```{r}
names(l_final_tl3)
```

The xlsx writer function is tested with

```{r}
write_tl_xlsx(pl_toplist = l_final_tl3, ptbl_trait = tbl_trait_name, ps_xlsx_file = "Toplisten_Stiere_CHbv_1904.xlsx")
```


# Additional Experimental Stuff
```{r}
names(l_final_tl)
```

```{r}
l_cur_df_list <- l_final_tl$BV
l_cur_df_list[s_cur_trait]
```


```{r}
l_final_tl$BV[s_cur_trait]
```

```{r}
l_final_tl[["BV"]][s_cur_trait]
```

```{r}
tbl_cur_tab <- l_final_tl[["BV"]][[s_cur_trait]]
tbl_cur_tab
```

```{r}
colnames(tbl_cur_tab)
```

